---
title: "Results.One"
author: "Ernesto Ramirez"
date: "January 27, 2016"
output:
  word_document:
       reference_docx: Background_v1.docx
       fig_caption: true
csl: apa-no-ampersand.csl
bibliography: library.bib
---
# Results
### Participants.
Participants were recruited between June and August 2015. Thirty-seven individuals responded to recruitment postings and expressed an interest in participating in the study. Of the 37 individuals who expressed interest, 33 completed the screening process for inclusion in the study. All 33 met the inclusion criteria and were asked to schedule the interview and connect their Fitbit data to Fitabase. Thirty-two individuals scheduled interviews and connected their Fitbit data to Fitabase, however two participants were not able to complete the interview due to time constraints. The Fitabase data connection was deleted for the two participants who were not able to complete the interview.  In total, 30 participants fully participated in the current study.

Table [TK] displays the demographic information for the sample that completed the study. The sample consisted of 11 (37%) men and 19 women (63%). Participants were between 23 and 60 years old, with a mean age of 36.60 years (SD = 10.56). The majority of participants reported that they were of Caucasian race/ethnicity (73%), were college educated (80% completed college or graduate degree), and worked full time (83%). Half the sample reported being married with another 40% reporting as “Single and Never Married.” Additionally, 33% of the sample reported a household income in excess of $100,000. Median income was $80,000 - $89,000 (one participant abstained from reporting household income information). Height and weight was self-reported and BMI was calculated. Participants had a BMI range of 17.28 to 59.67 kg/m2, with a mean of 25.85 kg/m<sup>2</sup> (SD = 8.47). Two participants were classified as “underweight” (6.67%, BMI < 25kg/m<sup>2</sup>), 24 participants (80%) were classified as “normal weight” (BMI = 25-30kg/m<sup>2</sup>), two participants were classified as "Obese Level I" (BMI = 30-35kg/m<sup>2</sup>), and two participants were classified as “Obese Level III" (BMI > 40kg/m<sup>2</sup>). Four participants reported membership in "Quantified Self" communities.

### Fitbit Devices
As previously mentioned, at the time of this study there were six different Fitbit devices available for purchase. Each of the six available devices were represented in the data set, with an additional device that is no longer available for purchase also represented (Ultra). In total, six participants were using the ChargeHR, six participants were using the Zip, five participants were using the Flex, five were using the One, two participants were using the Surge, and one participant was using the Ultra.

```{r}
library(knitr)
load("~/Dropbox/Dissertation.Data/Data/demographics.table.rdata")
kable(demographics.table, caption = "Table [TK]: Participants' Demographics and Characteristics")
```

### Fitbit Use, Wear Time, and Consistency
Participants in this study ranged from relatively new to using a Fitbit to long-term users, with a range of 98 to 2,014 days (*M* = 688.00, *SD* = 487.54). Initial review of all Fitbit data from the 30 participants in this study resulted in a total of 20,605 participant-days of available data. After completing the weartime processing steps described in the methods section a total of 15,941 valid participant-days of data were observed in the dataset. The ratio of valid days to total available days per participant ranged from 10.22% to 99.27%. The number of days of data at each step in the previously described analysis is presented in Table [TK]. A visual representation of all valid wear days is also presented in figure [TK].

```{r, echo=FALSE, results = 'asis'}
library(knitr)
library(plyr)
load("~/Dropbox/Dissertation.Data/FitbitData/ProcessedData/Weartime.rdata")
WearTable <- ddply(Daily.Activity.Final, c("id"), summarise,
                   'Initial Available Days'=length(id),
                   'Valid Days (MET Processing)'=sum(ValidDay.Counts),
                   'Valid Days (Censured)'=sum(ValidDay.Censured),
                   'Valid Days (Final)'=sum(Valid.Final)
)
WearTable$'Percent of Initial Available Days Included in Analysis' <- WearTable$`Valid Days (Final)` / WearTable$`Initial Available Days`
kable(WearTable, caption = "Table [TK]")
```

```{r, echo=FALSE, fig.cap="Figure TK. Visualization of Valid Fitbit Wear per Participant"}
ggplot(Weartime, aes(x=Date, y=Valid.Final)) +
  geom_bar(stat="identity") +
  scale_y_continuous(breaks=NULL, labels = NULL, name = "") +
  facet_grid(id ~.)
```
On average participants wore their Fitbit and recorded a valid day for 20.70 consecutive days (*SD* = 48.17). Values for the minimum, maximum, mean, and standard deviation of consecutive day streaks for valid and non-wear days are included in Table [TK]. The longest consecutive streak of valid wear time lasted nearly two years (714 days, Participant 17). The longest consecutive streak of non-wear time was 597 days (Participant 18).

```{r}
library(knitr, echo=FALSE)
load("~/Dropbox/Dissertation.Data/Data/StreakTable.rdata")
kable(StreakTable, caption = "Table [TK]: Consecutive Non-Wear and Valid Wear Days per Participant")
```
## Physical Activity Outcome Variables

### Steps
For all valid days, participants in the current study totaled 154,546,189 steps. The mean across all participants was 9,694.87 steps per day (*SD* = 5,309.45). Steps per valid day ranged from a 109 to 68,565 steps. Descriptive data on steps per valid day for each participant is shown in Table [TK]. Participant step counts on valid days are visualized in figures TK - TK.

```{r, echo=FALSE}
library(knitr)
load("~/Dropbox/Dissertation.Data/Data/StepDescriptives.Valid.Table.rdata")
kable(StepDescriptives.Valid.Table, caption = "Table [TK]: Physical Activity Descriptives: Steps", row.names = FALSE)
```

### MVPA Bout Minutes
For all valid days, participants in the current study totaled 529,078 minutes of MVPA that were performed in bouts of at least 10 minutes. The mean amount of MVPA Bout Minutes was 33.19 minutes
per day (*SD* = 42.70). Per participant minutes of MVPA, when classified in bouts, ranged from 0 to 474 minutes per day. Half of all participants did not obtain a mean of at least 30 minutes of MVPA per day for observed valid days.

```{r, echo=FALSE}
library(knitr)
load("~/Dropbox/Dissertation.Data/Data/MVPABoutsDescriptives.Valid.Table.rdata")
kable(MVPABoutsDescriptives.Valid.Table.rdata, caption = "Table [TK]: MVPA Bout Minutes Summary Statistics per Participant", row.names = FALSE)
